Staged Abstract Interpreters

### 2019-09-08 paper/20_prelim.tex
typos

\paragraph{Monadic Operations} We define the return type of the interpreter
@Ans@, as a monadic type @AnsM[_]@ wrapping the type @Value@.
As mentioned in Section ~\ref{monadscala}, to use the @for@-comprehension
syntax, certain operations have to be added on the type @AnsM@. Here, we use a
structural type @MonadOps@ to require that type @AnsM@ should at least
implement @map@ and @flatMap@. It is worth noting that @MonadOps@ takes another
type parameter @R[_]@ as the binding-time, i.e., @MonadOps[R, AnsM, T]@.
Inside of @MonadOps@, @R[_]@ annotates the data types @A@ and @B@ that are
encapsulated by the monad, but not the monad type @M@ itself. When the generic
interpreter acts as a compiler, we will replace the monads with the ones that
work on staged data values.

We also declare several operations to manipulate environments and
stores. These methods return monadic values of type @AnsM[_]@, which
may be parameterized over the environment or store type, or merely a @Unit@
value for performing effects. For example, @local_env@ installs a new environment
@ρ@ when evaluating the monadic value @ans@; @set_store@ takes a pair of
addresses and values (potentially staged) and updates the store accordingly.

@@

\paragraph{Primitive Operations} Next, we define a few primitive operations.
First, we delcare two versions of @alloc@. The first one takes a store and an
identifier and produces a fresh address of non-monadic type @R[Addr]@. Since
the freshness of the address may depend on the store, which might be a
next-stage value as indicated by its type, the type of addresses is
consequently wrapped by @R[_]@. The other @alloc@ is simply the monadic version
of the addresses, and therefore can be used in monadic computations.
An auxiliary method @get@ retrieves the value of an identifier @x@, through the
environment and store.
\begin{lstlisting}
  def alloc(σ: R[Store], x: Ident): R[Addr];  def alloc(x: Ident): AnsM[Addr]
  def get(σ: R[Store], ρ: R[Env], x: Ident): R[Value]
\end{lstlisting}

Other primitive operations in the interpreter handle the language constructs.  The
methods @num@ and @close@ deal with primitive values, which lift literal terms
(e.g., lambdas) to our value representation (e.g., closures).
Conditionals and arithmetic are handled by @br0@ and @arith@, respectively. The
method @ap_clo@ is used for applying functions, which takes a function value
and an argument value. Note that the @Env@, @Store@, and @Value@ are all
annotated by @R[_]@, because they are potentially next-stage values when the
interpreter acts as a compiler.
\begin{lstlisting}
  def num(i: Int): Ans
  def close(ev: Expr => Ans)(λ: Lam, ρ: R[Env]): R[Value]
  def br0(test: R[Value], thn: => Ans, els: => Ans): Ans
  def arith(op: Symbol, v1: R[Value], v2: R[Value]): R[Value]
  def ap_clo(ev: Expr => Ans)(fun: R[Value], arg: R[Value]): Ans
\end{lstlisting}

>>>>

\paragraph{Monadic Operations} We define the return type of the interpreter
@Ans@ as a monadic type @AnsM[_]@ wrapping the type @Value@.
As mentioned in Section~\ref{monadscala}, in order to use the @for@-comprehension
syntax certain operations must be added on the type @AnsM@. Here, we use a
structural type @MonadOps@ to require that type @AnsM@ must (at least)
implement @map@ and @flatMap@. It is worth noting that @MonadOps@ takes another
type parameter @R[_]@ as the binding-time, i.e., @MonadOps[R, AnsM, T]@.
Inside of @MonadOps@, @R[_]@ annotates the data types @A@ and @B@ that are
encapsulated by the monad, but not the monad type @M@ itself. When the generic
interpreter acts as a compiler, we will replace the monads with the ones that
work on staged data values.

We also declare several operations to manipulate environments and
stores. These methods return monadic values of type @AnsM[_]@, which
may be parameterized over the environment or store type, or merely a @Unit@
value for performing effects. For example, @local_env@ installs a new environment
@ρ@ when evaluating the monadic value @ans@; @set_store@ takes a pair of
addresses and (potentially staged) values, and updates the store accordingly.

@@

\paragraph{Primitive Operations} Next, we define a few primitive operations.
First, we declare two versions of @alloc@. The first takes a store and an
identifier and produces a fresh address of non-monadic type @R[Addr]@. Since
the freshness of the address may depend on the store, which might be a
next-stage value as indicated by its type, the type of addresses is
consequently wrapped by @R[_]@. The other @alloc@ is simply the monadic version
of the addresses, and can therefore be used in monadic computations.
An auxiliary method @get@ retrieves the value of an identifier @x@ through the
environment and store.
\begin{lstlisting}
  def alloc(σ: R[Store], x: Ident): R[Addr];  def alloc(x: Ident): AnsM[Addr]
  def get(σ: R[Store], ρ: R[Env], x: Ident): R[Value]
\end{lstlisting}

Other primitive operations in the interpreter handle the language constructs.  The
methods @num@ and @close@ deal with primitive values, which lift literal terms
(e.g., lambdas) to our value representation (e.g., closures).
Conditionals and arithmetic are handled by @br0@ and @arith@, respectively. The
method @ap_clo@ is used for applying functions, which takes a function value
and an argument value. Note that the @Env@, @Store@, and @Value@ are all
annotated by @R[_]@, as they are potentially next-stage values when the
interpreter acts as a compiler.
\begin{lstlisting}
  def num(i: Int): Ans
  def close(ev: Expr => Ans)(λ: Lam, ρ: R[Env]): R[Value]
  def br0(test: R[Value], thn: => Ans, els: => Ans): Ans
  def arith(op: Symbol, v1: R[Value], v2: R[Value]): R[Value]
  def ap_clo(ev: Expr => Ans)(fun: R[Value], arg: R[Value]): Ans
\end{lstlisting}
